# Load packages and data
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)

k <- read_xlsx("data/Khapra Rice Experiment Combined.xlsx") %>%
  select(samp = 'Sample Name', level = 3, ct = 5, quant = 8) %>%
  arrange(samp) %>%
  # remove an odd redo of a negative sample
  filter(samp != "1/8D Redo") %>%
  group_by(samp) %>%
  mutate(pcrnum = LETTERS[1:3]) %>%
  pivot_wider(id_cols = c(1,2), values_from = 3, names_from = pcrnum) %>%
  ungroup() %>%
  mutate(hits = 1-is.na(as.numeric(A)) + 
           1-is.na(as.numeric(B)) + 
           1-is.na(as.numeric(C)), 
         subs = 3,
        dna = case_when(level == "1/2" ~ 0.5,
                         level == "1/4" ~ 0.25,
                         level == "1/8" ~ 0.125,
                         level == "16" ~ 16,
                         level == "1" ~ 1,
                         level == "2" ~ 2,
                         level == "4" ~ 4,
                         level == "8" ~ 8),
        bd = case_when(level == "1/2" ~ 5.84,
                         level == "1/4" ~ 2.92,
                         level == "1/8" ~ 1.46,
                         level == "16" ~ 187,
                         level == "1" ~ 11.69,
                         level == "2" ~ 23.38,
                         level == "4" ~ 46.75,
                         level == "8" ~ 93.5),
        pres = (hits>0)*1)

# Data formatting for qPCR-replicate-based multilevel occupancy modeling in JAGS

# Standardize covariates
ldens <- log(k$dna)
ldensS <- wiqid::standardize(ldens)

# Create values of density to predict
ldens.pred = log(seq(0.089,16, length.out = 500))
ldens.predS = wiqid::standardize2match(ldens.pred, ldens)
# 0.125 = 1.46 beetles
# 0.089 = 1 beetle (note: original estimate was 0.0855)
lonebeetle.pred <- log(0.089)
lonebeetle.predS <- wiqid::standardize2match(log(0.089), ldens)

# Bundle and summarize data set for JAGS
khapra.jags.data <- list(y = k$hits, 
                         n = k$subs,
                         y01 = (k$hits>0)*1,
                           n.samp = nrow(k), 
                           dens = ldens,
                           densS = ldens, # ldensS to use z-score density
                           dens.predS = ldens.pred,
                           onebeetle.predS = lonebeetle.pred, # lonebeetle.predS for z-score
                         npred = length(ldens.pred))
```
# Define Bayesian qPCR-replicate-based multilevel occupancy model in JAGS
JAGS code modified from Kéry & Royle (2015, pp. 606-607).

Kéry, M., & Royle, J. A. (2015). Applied Hierarchical Modeling in Ecology: Analysis of distribution, abundance and species richness in R and BUGS: Volume 1: Prelude and Static Models. London: Academic Press. 
```{r}
cat(file = "scripts/khapra_PCR.txt", 
    "
model {

# Priors and model for params
int.p ~ dunif(0,1)           # Intercept detection probability
alpha0 <- logit(int.p)         # same on logit scale
beta.dens ~ dnorm(0, 0.01)   # density slope on logit(p)

# Likelihood
for (i in 1:n.samp){
   # detection process for PCR replicate k
   y[i] ~ dbin (p[i], n[i])
   logit(p[i]) <- alpha0 + beta.dens * densS[i]
     }

# Derived quantities
# Equation for 'at least one hit' for 3 PCR tech reps
for(i in 1:npred){
p.PCRrep3[i] <- 1-(1-ilogit(alpha0 + beta.dens*dens.predS[i]))^3
}

# Equation for 'at least one hit' for 1 PCR tech reps
for(i in 1:npred){
p.PCRrep1[i] <- ilogit(alpha0 + beta.dens*dens.predS[i])
}

# Equation for 'at least one hit' for increasing PCR tech reps
for(i in 1:6){
p.PCRrep.low[i] <- 1-(1-ilogit(alpha0 + beta.dens*onebeetle.predS))^i
}


}
",fill=TRUE)
```

# Run Bayesian qPCR-replicate-based multilevel occupancy model in JAGS
0.125 = 1.46 beetles
.089 = 1 beetle
```{r}
# Initial values
inits <- function() list(int.psi = runif(1), 
                         int.theta = runif(1),
                         int.p = runif(1), 
                         beta.dens = rnorm(1, 0, 1))

# Parameters monitored
params <- c("alpha0", "beta.dens", 
             "p.PCRrep1", "p.PCRrep.low", "p.PCRrep3")

#####################################
#### Run jags model
#####################################
khapra_PCR <- jagsUI::jags(data = khapra.jags.data,
                         inits = inits,
                         parameters.to.save = params,
                         model.file = "scripts/khapra_PCR.txt",
                         n.chains = 3,
                         # n.adapt = 100000,
                         n.burnin = 100000,
                         n.iter = 200000, 
                         n.thin = 20,
                         parallel = T,
                         verbose = TRUE,
                         modules = NULL)
khapra_PCR
saveRDS(khapra_PCR, "output/khapra_PCR_jags.rds")

khapra_PCR <- readRDS("output/khapra_PCR_jags.rds")

# what is the certainty of detecting 1 beetle at 6 technical replicates
khapra_PCR$mean$p.PCRrep.low # 0.9648582
khapra_PCR$q2.5$p.PCRrep.low # 0.8293482
khapra_PCR$q97.5$p.PCRrep.low # 0.9994474

# what is the certainty of detecting 1 beetle at 3 technical replicates
khapra_PCR$mean$p.PCRrep.low # 0.8429532
khapra_PCR$q2.5$p.PCRrep.low # 0.5868998
khapra_PCR$q97.5$p.PCRrep.low # 0.9764933

# get non-standardized betas
khapra_PCR$mean$alpha0 # 4.391
khapra_PCR$q2.5$alpha0 # 2.987798
khapra_PCR$q97.5$alpha0 # 6.316081
khapra_PCR$mean$beta.dens # 1.838996
khapra_PCR$q2.5$beta.dens # 1.002756
khapra_PCR$q97.5$beta.dens # 2.907368

khapra_PCR$mean$p.PCRrep3

# Examine model output
khapra_PCR

# Examine traceplots
# jagsUI::traceplot(khapra_PCR)
```
# Create Fig. 1
Graph detection probability of khapra beetle with density (1A) and also detection probability at 1 beetle density with increasing numbers of technical replicates (1B).
```{r}
# load model output 
# note: based on z-standardized log(grains)
    # so beta coefficients differ slightly from those reported
    # the non-standardized version is linked to above
    # both model converged, but it took multiple runs to do so
khapra_PCR <- readRDS("output/khapra_PCR_jags.rds")

# Format data for figure 1A
pv_all1 <- cbind.data.frame(
dna = ldens.pred,
med = apply(khapra_PCR$sims.list$p.PCRrep1, 2, mean),
q2.5 = apply(khapra_PCR$sims.list$p.PCRrep1, 2, 
      function(x) quantile(x, 0.025)),
q97.5 = apply(khapra_PCR$sims.list$p.PCRrep1, 2, 
      function(x) quantile(x, 0.975)),
q10 = apply(khapra_PCR$sims.list$p.PCRrep1, 2, 
      function(x) quantile(x, 0.1)),
q90 = apply(khapra_PCR$sims.list$p.PCRrep1, 2, 
      function(x) quantile(x, 0.90))
) %>%
  mutate(tech = "1 tech. rep.")

pv_all3 <- cbind.data.frame(
dna = ldens.pred,
med = apply(khapra_PCR$sims.list$p.PCRrep3, 2, mean),
q2.5 = apply(khapra_PCR$sims.list$p.PCRrep3, 2, 
      function(x) quantile(x, 0.025)),
q97.5 = apply(khapra_PCR$sims.list$p.PCRrep3, 2, 
      function(x) quantile(x, 0.975)),
q10 = apply(khapra_PCR$sims.list$p.PCRrep3, 2, 
      function(x) quantile(x, 0.1)),
q90 = apply(khapra_PCR$sims.list$p.PCRrep3, 2, 
      function(x) quantile(x, 0.90))
) %>%
  mutate(tech = "3 tech. rep.")

pv_all <- pv_all1 %>%
  bind_rows(pv_all3)

pt_est1 <- k %>%
  group_by(level, dna) %>%
  summarize(pct = mean(hits/subs)) %>%
  mutate(tech = "1 tech. rep.")

pt_est3 <- k %>%
  group_by(level, dna) %>%
  summarize(pct = mean(pres)) %>%
  mutate(tech = "3 tech. rep.")

pt_est <- pt_est1 %>%
  bind_rows(pt_est3)

# create "number of beetles" df
(nb <- data.frame(
  num = c(1,5,10,50,100),
  dna = c(1,5,10,50,100)*.089,
  y = 0.25))

(nb <- data.frame(
  num = c(1,2, 5,10,20, 50,100),
  dna = c(1,2, 5,10,20, 50,100)*.089,
  y = 0.25))

# Create figure 1A
(densplot <- pv_all %>%
  ggplot() +
  geom_abline(slope = 0, intercept = 0.95, 
              color = "darkgray", size=1, linetype = 2) +
  geom_ribbon(aes(x = exp(dna), ymin = q2.5, ymax = q97.5, 
                  fill = tech), 
              alpha = 0.5, color = "transparent") +
    geom_ribbon(aes(x = exp(dna), ymin = q10, ymax = q90, 
                  fill = tech),
              alpha = 0.5, color = "transparent") +
  scale_fill_manual(values = c("firebrick", "steelblue")) +
  geom_line(aes(x = exp(dna), y = med, group = tech), size = 1) +
  geom_point(aes(x = dna, y = pct), 
             size = 3, color = "black",
             data = pt_est1) +
  scale_color_manual(values = c("black", "black")) +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15)) +
  labs(y = "Detection probability", 
       x = "Infested grains / 100 g clean rice") +
  annotate("text", x = 1.2, y = 0.335, label = "Number of beetles",
           color = "firebrick") +
  theme(strip.text = element_text(
    size = 15,
    colour = "black",
    angle = 0
  )) +
  theme(legend.position = "") +
  geom_rug(aes(x = dna), data = nb, color = "firebrick") +
  geom_text(aes(x = dna, y = y+0.02, label = num), 
            data = nb, color = "firebrick", size = 4) +
  scale_x_log10(breaks = c(0.125, 0.25, 0.5, 1, 2, 4, 8, 16),
                labels = c("1/8", "1/4", "1/2", "1", 
                           "2", "4", "8", "16"))
)

# ggsave("figures/Fig_khapra_detection_density.png", 
#        width = 4, height = 4, dpi = 600)

# format data for plot 1B
pv_low <- cbind.data.frame(
reps = 1:6,
med = apply(khapra_PCR$sims.list$p.PCRrep.low, 2, mean),
q2.5 = apply(khapra_PCR$sims.list$p.PCRrep.low, 2, 
      function(x) quantile(x, 0.025)),
q97.5 = apply(khapra_PCR$sims.list$p.PCRrep.low, 2, 
      function(x) quantile(x, 0.975)),
q10 = apply(khapra_PCR$sims.list$p.PCRrep.low, 2, 
      function(x) quantile(x, 0.1)),
q90 = apply(khapra_PCR$sims.list$p.PCRrep.low, 2, 
      function(x) quantile(x, 0.90))
)

# Create figure 1B
(beetle1 <- pv_low %>%
  ggplot() +
  geom_abline(slope = 0, intercept = 0.95, 
              color = "darkgray", size=1, linetype = 2) +
  geom_errorbar(aes(x = reps, ymin = q2.5, ymax = q97.5),
              width = 0, size = .5) +
    geom_errorbar(aes(x = reps, ymin = q10, ymax = q90),
              width = 0, size = 1) +
  geom_line(aes(x = reps, y = med), size = 1) +
  geom_point(aes(x = reps, y = med), 
             size = 4, color = "black") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15)) +
  labs(y = "Detection probability at \n~ 1 beetle density", 
       x = "Number of qPCR technical replicates") +
  scale_x_continuous(breaks = c(1,2,3,4,5,6))
)

# ggsave("figures/Fig_khapra_1beetle_detection.png", 
#        width = 4, height = 4, dpi = 600)

# create combined figure
library(patchwork)

densplot + beetle1

ggsave("figures/Fig_khapra_detection_combined_400dpiC.png", 
       width = 8, height = 4, dpi = 400)
```