# Load packages and data
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)

k <- read_xlsx("data/Khapra Rice Experiment Combined.xlsx") %>%
  select(samp = 'Sample Name', level = 3, ct = 5, quant = 8) %>%
  arrange(samp) %>%
  # remove an odd redo of a negative sample
  filter(samp != "1/8D Redo") %>%
  group_by(samp) %>%
  mutate(pcrnum = LETTERS[1:3]) %>%
  pivot_wider(id_cols = c(1,2), values_from = 3, names_from = pcrnum) %>%
  ungroup() %>%
  mutate(hits = 1-is.na(as.numeric(A)) + 
           1-is.na(as.numeric(B)) + 
           1-is.na(as.numeric(C)), 
         subs = 3,
        dna = case_when(level == "1/2" ~ 0.5,
                         level == "1/4" ~ 0.25,
                         level == "1/8" ~ 0.125,
                         level == "16" ~ 16,
                         level == "1" ~ 1,
                         level == "2" ~ 2,
                         level == "4" ~ 4,
                         level == "8" ~ 8),
        bd = case_when(level == "1/2" ~ 5.84,
                         level == "1/4" ~ 2.92,
                         level == "1/8" ~ 1.46,
                         level == "16" ~ 187,
                         level == "1" ~ 11.69,
                         level == "2" ~ 23.38,
                         level == "4" ~ 46.75,
                         level == "8" ~ 93.5),
        pres = (hits>0)*1)

```

# Run model
```{r}
mod = glm(hits/subs ~ log(dna), family = "binomial", weights = subs,
    data = k)
summary(mod)

modbd = glm(hits/subs ~ bd, family = "binomial", weights = subs, 
    data = k)
summary(modbd)
```

# Plot model
```{r}
fit = predict(mod, 
              newdata = data.frame(dna = seq(0.125,16, 
                                                  length.out = 100)),
              se.fit = T)

pred = fit$fit
UL = fit$fit + 2*fit$se.fit
LL = fit$fit - 2*fit$se.fit

plotdata <- 
  data.frame(
dna = seq(0.125,16, length.out = 100),
fit1 = predict(mod, 
              newdata = data.frame(dna = seq(0.125,16, 
                                                  length.out = 100)),
              type = "response"),
fit = plogis(pred),
UL = plogis(UL),
LL = plogis(LL)
)

k %>%
  group_by(level) %>%
  summarize(pct = sum(hits)/24)
  
pt_est <- k %>%
  group_by(level, dna) %>%
  summarize(pct = mean(pres))

plotdata %>%
  ggplot() +
  geom_line(aes(x = dna, y = fit), size = 1) +
  geom_point(aes(x = dna, y = pct), data = pt_est, 
             color = "steelblue", size = 3, alpha = .7) +
    scale_x_continuous(breaks = c(0.125, 0.25, 0.5, 1, 2, 4, 8, 16),
  labels = c("", "", "", 1,2,4,8,16),
  limits = c(0,17)) +
  # scale_x_log10(breaks = c(),
  #               limits = c(0, 17)) +
  scale_y_continuous(limits = c(.8,1)) +
  labs(x = "Density of infested rice",
       y = "Detection probability") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

ggsave("figures/detection_prob_v_infested_rice_density.png",
       width = 6, height = 6, dpi = 400)

plotdata %>%
  ggplot() +
  geom_ribbon(aes(ymin = LL, ymax = UL, x = dna),
              fill = "steelblue", alpha = 0.5) +
  geom_line(aes(x = dna, y = fit), size = 1) +
  geom_point(aes(x = dna, y = pct), data = pt_est, 
             color = "steelblue", size = 3, alpha = .7) +
  #   scale_x_continuous(breaks = c(0.125, 0.25, 0.5, 1, 2, 4, 8, 16),
  # labels = c("", "", "", 1,2,4,8,16),
  # limits = c(0,17)) +
  scale_x_log10(limits = c(0.125, 17)) +
  scale_y_continuous(limits = c(.5,1)) +
  labs(x = "Density of infested rice\n(number of grains)",
       y = "Detection probability") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

ggsave("figures/detection_prob_v_infested_rice_density_log_CI.png",
       width = 6, height = 6, dpi = 400)

predict(mod, newdata = data.frame(dna = 0.0855), type = "response",
        se.fit = T)
# probabilty, 0.73, SE 0.14

```


# Data formatting for qPCR-replicate-based multilevel occupancy modeling in JAGS
```{r}

# Standardize covariates
ldens <- log(k$dna)
ldensS <- wiqid::standardize(ldens)

# Create values of density to predict
ldens.pred = log(seq(0.125,16, length.out = 100))
ldens.predS = wiqid::standardize2match(ldens.pred, ldens)
# 0.125 = 1.46 beetles
# 0.0855 = 1 beetle
lonebeetle.predS = wiqid::standardize2match(log(0.0855), ldens)

# Bundle and summarize data set
khapra.jags.data <- list(y = k$hits, 
                         n = k$subs,
                         y01 = (k$hits>0)*1,
                           n.samp = nrow(k), 
                           dens = ldens,
                           densS = ldensS,
                           dens.predS = ldens.predS,
                           onebeetle.predS = lonebeetle.predS)
```
# Define Bayesian qPCR-replicate-based multilevel occupancy model in JAGS
This is the primary analysis used in the manuscript, which treats the three 'suspected false positives' as negatives. JAGS code modified from Kéry & Royle (2015, pp. 606-607.

Kéry, M., & Royle, J. A. (2015). Applied Hierarchical Modeling in Ecology: Analysis of distribution, abundance and species richness in R and BUGS: Volume 1: Prelude and Static Models. London: Academic Press. 
```{r}
sink("khapra_PCR.txt")
cat("
model {

# Priors and model for params
int.p ~ dunif(0,1)           # Intercept detection probability
alpha0 <- logit(int.p)         # same on logit scale
beta.dens ~ dnorm(0, 0.01)   # density slope on logit(p)

# Likelihood
for (i in 1:n.samp){
   # detection process for PCR replicate k
   y[i] ~ dbin (p[i], n[i])
   logit(p[i]) <- alpha0 + beta.dens * densS[i]
     }

# Derived quantities
# Equation for 'at least one hit' for 3 PCR tech reps
for(i in 1:100){
p.PCRrep3[i] <- 1-(1-ilogit(alpha0 + beta.dens*dens.predS[i]))^3
}

# Equation for 'at least one hit' for 1 PCR tech reps
for(i in 1:100){
p.PCRrep1[i] <- ilogit(alpha0 + beta.dens*dens.predS[i])
}

# Equation for 'at least one hit' for increasing PCR tech reps
for(i in 1:6){
p.PCRrep.low[i] <- 1-(1-ilogit(alpha0 + beta.dens*onebeetle.predS))^i
}


}
",fill=TRUE)
sink()
```

# Run Bayesian qPCR-replicate-based multilevel occupancy model in JAGS
0.125 = 1.46 beetles
.0855 = 1 beetle
```{r}
# Initial values
inits <- function() list(int.psi = runif(1), 
                         int.theta = runif(1),
                         int.p = runif(1), 
                         beta.dens = rnorm(1, 0, 1))

# Parameters monitored
params <- c("alpha0", "beta.dens", 
             "p.PCRrep1", "p.PCRrep.low", "p.PCRrep3")

#####################################
#### Run jags model
#####################################
khapra_PCR <- jagsUI::jags(data = khapra.jags.data,
                         inits = inits,
                         parameters.to.save = params,
                         model.file = "khapra_PCR.txt",
                         n.chains = 3,
                         n.adapt = NULL,
                         n.burnin = 10000,
                         n.iter = 20000, 
                         n.thin = 10,
                         parallel = T,
                         verbose = TRUE,
                         modules = NULL)

khapra_PCR_bern <- jagsUI::jags(data = khapra.jags.data,
                         inits = inits,
                         parameters.to.save = params,
                         model.file = "khapra_PCR_bern.txt",
                         n.chains = 3,
                         n.adapt = NULL,
                         n.burnin = 10000,
                         n.iter = 20000, 
                         n.thin = 10,
                         parallel = T,
                         verbose = TRUE,
                         modules = NULL)

# Examine model output
# khapra_PCR
khapra_PCR_bern


# Examine traceplots
# jagsUI::traceplot(NJ_vineyards_PCR)
```
# Create Fig.
Graph cumulative probability of detection in occupied vineyards with increasing survey effort. This is the primary analysis used in the manuscript, which treats the three 'suspected false positives' as negatives.
```{r}
# Format data and create figure
# Note: you'll need to load the sum_posts function in the previous chunk first
options(digits = 10)
pv_all1 <- cbind.data.frame(
dna = ldens.pred,
med = apply(khapra_PCR$sims.list$p.PCRrep1, 2, mean),
q2.5 = apply(khapra_PCR$sims.list$p.PCRrep1, 2, 
      function(x) quantile(x, 0.025)),
q97.5 = apply(khapra_PCR$sims.list$p.PCRrep1, 2, 
      function(x) quantile(x, 0.975)),
q10 = apply(khapra_PCR$sims.list$p.PCRrep1, 2, 
      function(x) quantile(x, 0.1)),
q90 = apply(khapra_PCR$sims.list$p.PCRrep1, 2, 
      function(x) quantile(x, 0.90))
) %>%
  mutate(tech = "1 tech. rep.")

pv_all3 <- cbind.data.frame(
dna = ldens.pred,
med = apply(khapra_PCR$sims.list$p.PCRrep3, 2, mean),
q2.5 = apply(khapra_PCR$sims.list$p.PCRrep3, 2, 
      function(x) quantile(x, 0.025)),
q97.5 = apply(khapra_PCR$sims.list$p.PCRrep3, 2, 
      function(x) quantile(x, 0.975)),
q10 = apply(khapra_PCR$sims.list$p.PCRrep3, 2, 
      function(x) quantile(x, 0.1)),
q90 = apply(khapra_PCR$sims.list$p.PCRrep3, 2, 
      function(x) quantile(x, 0.90))
) %>%
  mutate(tech = "3 tech. rep.")

pv_all <- pv_all1 %>%
  bind_rows(pv_all3)

pt_est1 <- k %>%
  group_by(level, dna) %>%
  summarize(pct = mean(hits/subs)) %>%
  mutate(tech = "1 tech. rep.")

pt_est3 <- k %>%
  group_by(level, dna) %>%
  summarize(pct = mean(pres)) %>%
  mutate(tech = "3 tech. rep.")

pt_est <- pt_est1 %>%
  bind_rows(pt_est3)

options(digits = 2)
pv_all %>%
  ggplot() +
  geom_ribbon(aes(x = exp(dna), ymin = q2.5, ymax = q97.5, group = 1), 
              alpha = 0.5, color = "transparent") +
    geom_ribbon(aes(x = exp(dna), ymin = q10, ymax = q90, group = 1),
              alpha = 0.5, color = "transparent") +
  scale_fill_manual(values = c("firebrick", "darkgray")) +
  geom_line(aes(x = exp(dna), y = med, group = 1), size = 1.5) +
  geom_point(aes(x = dna, y = pct), color = "black", 
             size = 2, data = pt_est) +
  scale_color_manual(values = c("black", "black")) +
  facet_wrap(~tech) +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15)) +
  labs(y = "Detection probability", 
       x = "Density of infested rice") +
  theme(strip.text = element_text(
    size = 15,
    colour = "black",
    angle = 0
  )) +
  theme(legend.position = "") +
  geom_abline(slope = 0, intercept = 0.95, 
              color = "darkgray", size=1, linetype = 2) +
  scale_x_log10(breaks = c(0.125, 0.25, 0.5, 1, 2, 4, 8, 16))
  # scale_x_continuous(breaks = c(0.125, 0.25, 0.5, 1, 2, 4, 8, 16),
  # labels = c("", "", "", 1,2,4,8,16) +
  # limits = c(0,17) #+
  # scale_y_continuous(limits = c(0.8,1))

# ggsave("Fig_2_vineyard_cumulative_detection_Bayes.pdf", width = 7, height = 6, dpi = 600, device=cairo_pdf)
```



# Define Bayesian qPCR-replicate-based multilevel occupancy model in JAGS
This is the primary analysis used in the manuscript, which treats the three 'suspected false positives' as negatives. JAGS code modified from Kéry & Royle (2015, pp. 606-607.

Kéry, M., & Royle, J. A. (2015). Applied Hierarchical Modeling in Ecology: Analysis of distribution, abundance and species richness in R and BUGS: Volume 1: Prelude and Static Models. London: Academic Press. 
```{r}
sink("khapra_PCR2.txt")
cat("
model {

# Priors and model for params
# int.psi ~ dunif(0,1)         # Intercept of capture probability
# gamma0 <- logit(int.psi)       # same on logit scale

int.theta ~ dunif(0,1)       # Intercept capture probability
beta0 <- logit(int.theta)      # same on logit scale
beta.dens ~ dnorm(0, 0.1)   # Survey round covariate on logit(plot-level occupancy)

int.p ~ dunif(0,1)           # Intercept detection probability
alpha0 <- logit(int.p)         # same on logit scale

# Likelihood
for (i in 1:n.vineyard){
   # Occurrence in vineyard i
   z[i] ~ dbern(psi[i])
   logit(psi[i]) <- gamma0
   for (j in 1:n.plots){
      # Occurrence in plot sample j
      a[i,j] ~ dbern(z[i] * theta[i,j])
       logit(theta[i,j]) <- beta0 + beta.round * round[i,j]
      for (k in 1:n.PCRrep){
         # detection process for PCR replicate k
         y[i,j,k] ~ dbern(a[i,j] * p[i,j,k])
         logit(p[i,j,k]) <- alpha0
      }
   }
 }

# Derived quantities
# Vineyard-level occupancy
logit(psi.vineyards) <- gamma0
# Theta or eDNA capture probability in round 1
logit(theta.round1) <- beta0 + beta.round
# Theta or eDNA capture probability in round 2
logit(theta.round2) <- beta0
# Detection probabity for an individual qPCR replicate
logit(p.PCRrep) <- alpha0

# Equation for 'at least one hit' for PCR
for(i in 1:10){
p.PCRrep_cum[i] <- 1-(1-p.PCRrep)^i
}

}
",fill=TRUE)
sink()
```

# Same but Bernoulli
```{r}
sink("khapra_PCR_bern.txt")
cat("
model {

# Priors and model for params
int.p ~ dunif(0,1)           # Intercept detection probability
alpha0 <- logit(int.p)         # same on logit scale
beta.dens ~ dnorm(0, 5)   # density slope on logit(p)

# Likelihood
for (i in 1:n.samp){
   # detection process for PCR replicate k
   y01[i] ~ dbern(p[i])
   logit(p[i]) <- alpha0 + beta.dens * densS[i]
     }

# Derived quantities
# Equation for 'at least one hit' for 3 PCR tech reps
for(i in 1:100){
p.PCRrep3[i] <- 1-(1-ilogit(alpha0 + beta.dens*dens.predS[i]))^3
}

# Equation for 'at least one hit' for 1 PCR tech reps
for(i in 1:100){
p.PCRrep1[i] <- 1-(1-ilogit(alpha0 + beta.dens*dens.predS[i]))^1
}

# Equation for 'at least one hit' for increasing PCR tech reps
for(i in 1:6){
p.PCRrep.low[i] <- 1-(1-ilogit(alpha0 + beta.dens*-0.7397835))^i
}


}
",fill=TRUE)
sink()
```
# plot

```{r}



```




